---
title: "ggplot2_workshop"
output: html_document
date: "2023-09-18"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install and Load Packages

Install and load the following packages: tidyverse (includes ggplot and dplyr) & ggrepel

```{r}
library(tidyverse)
library(ggrepel)
```

```{r Answer}
# Install packages. You only need to do this once on a given machine
#install.packages('tidyverse', 'ggrepel') 

# Load packages. Everytime you open a new workspace.
library(tidyverse)
library(ggrepel)

```

## Read in Data

1.  set working directory
2.  Read in happiness.csv file and name the data frame "happy"
3.  Take a look at data using str(),
4.  change the data class for country and continent

```{r pressure, echo=FALSE}

setwd("/Users/lmn7/Documents/workshops/ggplot2Workshop/Rforfun")

happy = read.csv("happiness.csv", head=T)

str(happy)

happy$country_name <- as.factor(happy$country_name)
happy$continent <- as.factor(happy$continent)
```

```{r Answer}

#you will need to direct this to your own folder using the file path. Or by going to the "session" menu, selecting "set working directory" and navigating to the appropriate folder
setwd("/Users/lmn7/Downloads/ggplot2_workshop-main")

happy = read.csv("Happiness.csv")
str(happy)

happy$country_name <- as.factor(happy$country_name)
happy$continent <- as.factor(happy$continent)
```

## Make a boxplot

1.  Use geom_boxplot() to plot the median & standard error for happiness ratings within each continent
2.  Add individual points on top of the boxplot using geom_point()
3.  jitter the points using geom_jitter()
4.  Make all points dark blue (using col='darkblue') and semi-transparent (using alpha=0.5)
5.  remove the outlier from the geom_box plot now that you have all points included on the figure by specifying outlier.shape=NA

```{r}

```

```{r Answer }
ggplot(happy, aes(continent, happiness))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(alpha=.5, col='darkblue')

```

## Sort the data

1.  Reorder the boxplots above in descending order (hint: specify the x axis using the reorder() function and specifying "FUN = mean" within the function)

```{r}

```

```{r Answer}
ggplot(happy, 
       aes(reorder(continent, -happiness, FUN=median), happiness))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(alpha=.5, col='darkblue')
```

## Make a Scatterplot: manipulate data

1.  Plot GDP vs happiness for all country and years
2.  GDP in the dataset has been log transformed. Remove the log transformation using the exp() function so that the figure plots the actual dollar values
3.  Subset the data so that you are only plotting observations from 2019

```{r}

```

```{r Answer}
ggplot(happy %>% filter(year==2019), 
       aes(exp(GDP), happiness))+
  geom_point()
```

## Adjusting labels and titles

1.  Using labs() layer:

    A. update the x and y labels to read: "GDP per capita" and "Average happiness"

    B. Add the following title and subtitle: title= 'Richer countries tend to be happier countries' subtitle ='Though the richest countries arent necessairily the happiest countries (year 2019)'

```{r}



```

```{r Answer}
ggplot(happy %>% filter(year==2019), 
       aes(exp(GDP), happiness))+
  geom_point() +
  labs(x='GDP per capita', y='Average happiness',title= 'Richer countries tend to be happier countries', subtitle='Though the richest countries arent necessairily the happiest countries')
```

## Conditional Formatting and Scales

1.  Amend the code below so that the countries with a happiness rating above 7.5 are a different color
2.  Change the colors so that the points above 7.5 are orange, and those below 7.5 are black. Make all points semi-transparent.

```{r}



```

-   Note: There are multiple ways to do this. Two approaches:

    1: Set conditional formatting within a single layer. ggplot will use the standard colors, so to get orange and black, you need to create a list of the colors you want, and then specify that you want ggplot to use those colors using scale_colour_manual(values = yourcolorlist)

-   2: Subset the data and plot the black and orange points seperately as two seperate geoms

```{r Answer}
newColorPalette = c('black','orange')

ggplot(happy %>% filter(year==2019), 
       aes(exp(GDP), happiness))+
  geom_point(aes(col=happiness>7.5)) +
  labs(x='GDP per capita', y='Average happiness',title= 'Richer countries tend to be happier countries', subtitle='Though the richest countries arent necessairily the happiest countries')+
  scale_color_manual(values = newColorPalette)


```

## Text labels

Amend the code below to accomplish the following two goals:

1\. Plot only the data from 2019

2\. Label the countries that have happiness \>7.5 with the country name using the geom_text_repel() function from the extension package ggrepel

```{r}


```

```{r Answer}

newColorPalette = c('black','orange')

ggplot(happy %>% filter(year==2019), aes(exp(GDP), happiness))+
  geom_point(aes(col=happiness>7.5)) +
  labs(x='GDP per capita', y='Average happiness',title= 'Richer countries tend to be happier countries', subtitle='In 2019 the happiest countries were all Nordic countries')+
  scale_color_manual(values = newColorPalette)+
  geom_text_repel(data=happy %>% filter(year==2019 & happiness>7.5),
                  aes(x=exp(GDP), y=happiness,label=country_name),col='orange')


```
