---
title: "ggplot2_demo"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

## Install and Load Packages

Install and load the following packages: tidyverse (includes ggplot and dplyr) & ggrepel

```{r}
# Install packages. You only need to do this once on a given machine
#install.packages('tidyverse') 

# Load packages. Everytime you open a new workspace.
library(tidyverse)
library(ggrepel) #ggrepel is a useful package for formatting labels

```

## Read in Data

1.  set working directory
2.  Read in happiness.csv file and name the data frame "happy"

```{r}

happy <- read_csv("Happiness.csv")

happy

```

## Make a boxplot

1.  Use `geom_boxplot()` to plot the median & standard error for happiness ratings within each continent
2.  Add individual points on top of the boxplot using `geom_point()`
3.  jitter the points using `geom_jitter()`
4.  Make all points dark blue (using col='blue') and semi-transparent (using `alpha=0.5`)
5.  remove the outlier from the geom_box plot now that you have all points included on the figure by specifying `outlier.shape=NA`

```{r}



```

```{r Solution}

happy |> 
  ggplot(aes(continent, happiness))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(col='blue',alpha=0.5)


```

## Sort the data

1.  Reorder the boxplots in descending order (hint: specify the x axis using the `reorder()` function and specifying `FUN = median` within the function)
2.  subset the data to just year 2019

```{r}
happy |> 
  ggplot(aes(continent, happiness))+
  geom_boxplot()+
  geom_jitter()
```

```{r Solution}
happy |> 
  filter(year==2019) |> 
  ggplot(aes(reorder(continent,happiness, FUN = median), happiness))+
  geom_boxplot()+
  geom_jitter()



```

## Make a Scatterplot: manipulate data

1.  Plot GDP vs happiness for all country and years
2.  GDP in the dataset has been log10 transformed. Remove the log transformation using the `exp()` function so that the figure plots the actual dollar values
3.  Subset the data so that you are only plotting observations from 2019.

```{r}

happy |> 
  filter(year==2019) |> 
  ggplot(aes(____, _____))+
  geom_point()

?exp

```

```{r Solution}

happy |> 
  filter(year==2019) |> 
  ggplot(aes(exp(GDP), happiness))+
  geom_point()

```

## Adjusting labels and titles

1.  Using labs() layer:

    A. update the x and y labels to read: `"GDP per capita"` and `"Average happiness"`

    B. Add the following title and subtitle: `title= 'The happiest countries are also among the richest'` `subtitle ='In 2019 the happiest countries were all Nordic countries'`

2.  Use `theme_minimal()` to change the theme so that the background is white instead of grey

```{r}
happy |>  
  filter(year==2019) |> 
  ggplot(aes(exp(GDP), happiness))+
  geom_point()

```

```{r Solution}

happy |>  
  filter(year==2019) |> 
  ggplot(aes(exp(GDP), happiness))+
  geom_point()+
  labs(title= 'The happiest countries are also among the richest',
       subtitle ='In 2019 the happiest countries were all Nordic countries',
       x="GDP per capita", y="Average happiness")+
  theme_minimal()


```

## Conditional Formatting and Scales

1.  Amend the code below so that the countries with a happiness rating above 7.5 are a different color

2.  Change the colors so that the points above 7.5 are `"darkorange"`, and those below 7.5 are `"darkgrey"`. Make all points semi-transparent using `alpha=0.7`

    Note: There are multiple ways to do this. Two approaches:

    1: Set conditional formatting within a single layer. ggplot will use the standard colors, so to get orange and black, you need to create a list of the colors you want, and then specify that you want ggplot to use those colors using `scale_colour_manual(values = yourcolorlist)`

    2: Subset the data and plot the black and orange points seperately as two seperate geoms

```{r}
ggplot(happy %>% filter(year==2019), 
       aes(exp(GDP), happiness))+
  geom_point()+
  labs(x='GDP per capita', y='Average happiness',
       title= 'The happiest countries are also among the richest', 
       subtitle='In 2019 the happiest countries were all Nordic countries')+
  theme_minimal()

```

```{r solution}
prettyColors = c('darkgrey','darkorange')

ggplot(happy, aes(exp(GDP), happiness))+
  geom_point(aes(col=happiness>7.5), alpha=0.7) +
  labs(x='GDP per capita', y='Average happiness',
       title= 'The happiest countries are also among the richest', 
       subtitle='In 2019 the happiest countries were all Nordic countries')+
  theme_minimal()+
  scale_color_manual(values = prettyColors)
```

## Text labels

Amend the code below to accomplish the following two goals:

1\. Plot only the data from 2019

2\. Label the countries that have happiness \>7.5 with the country name using the `geom_text_repel()` function from the extension package ggrepel

```{r}
prettyColors = c('darkgrey','darkorange')

happy |> 
  filter(year==2019) |> 
  ggplot(aes(exp(GDP), happiness))+
  geom_point(aes(col=happiness>7.5), alpha=0.7) +
  labs(x='GDP per capita', y='Average happiness',
       title= 'The happiest countries are also among the richest', 
       subtitle='In 2019 the happiest countries were all Nordic countries')+
  theme_minimal()+
  scale_color_manual(values = prettyColors)
```

```{r solution}


prettyColors = c('darkgrey','darkorange')

happy |> 
  filter(year==2019) |> 
  ggplot(aes(exp(GDP), happiness))+
  geom_point(aes(col=happiness>7.5), 
             alpha=0.5, 
             show.legend=FALSE) +
  labs(x='GDP per capita', y='Average happiness',title= 'The happiest countries are also among the richest', subtitle='In 2019 the happiest countries were all Nordic countries')+
  theme_minimal()+
  scale_color_manual(values = prettyColors)+
  geom_text_repel(data=happy |> filter(happiness>7.5 & year == 2019),
                  aes(label=country_name), 
                  col='darkorange',
                  show.legend=FALSE)

```
